\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species

### 2024-03-06 ms.Rnw
modify Rnw and tex
Update

<<Setup, include=FALSE,echo=FALSE,results=hide>>=
 load('all_shared.RData')

parms = mif.estimate
parms = as.data.frame(t(parms))

@

@@

$K_F$ & Alga carrying capacity&$10^5 cells \cdot L ^{-1} $ & \Sexpr{mysignif(parms$K_F,2)}\\
$r_n$ & Native growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$rn,2)}\\
$r_i$ & Invasive growth rate &$10^{-5} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$ri,2)}\\
$\theta_{I_n}$ & Native infected host mortality rate&$ day^{-1}$  & \Sexpr{mysignif(parms$theta_In,2)}\\
$\theta_{I_i}$ & Invasive infected host mortality rate&$ day^{-1}$  & \Sexpr{mysignif(parms$theta_Ii,2)}\\
$\theta_{S_i}$ & Invasive susceptible host mortality rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_Si,2)}\\
$\theta_{S_n}$ & Native susceptible host mortality rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_Sn,2)}\\
$\theta_{P}$ & Spore degradation rate&$ day^{-1}$ & \Sexpr{mysignif(parms$theta_P,2)}\\
$p_n$ &Native probability of infection per spore &$10^{-3} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probn,2)}\\
$p_i$ & Invasive probability of infection per spore&$10^{-3} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probi,2)}\\
$f_{S_n}$ & Native susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Sn,2)}\\
$f_{S_i}$ & Invasive susceptible host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_Si,2)} \\
$f_{I_n}$ & Native infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_In,2)} \\
$f_{I_i}$ & Invasive infected host filtering rate&$L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Ii,2)}\\

@@

$\alpha$ & Alga maximum exponential growth rate&$day ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
$\gamma_n$ & Native host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
$\gamma_i$ & Invsaive host consumption rate&$L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\

@@

$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
$\sigma_{F}$ & Standard deviation of brownian motion of alga&$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$\sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\

@@

\begin{align}
&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n}, \\
&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
&dF=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
&\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}^2dt  \right) \\
&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i}^2dt \right) 
\end{align}

>>>>

<<Results, include=FALSE,echo=FALSE,results=hide>>=
load('all_shared.RData')

parms = mif.estimate
# parms = as.data.frame(t(parms))

parms$K_F = parms$K_F * 1000
parms$ri = parms$ri * 10
parms$rn = parms$rn * 10
parms$theta_Ii = parms$theta_Ii * 10
parms$theta_In = parms$theta_In * 10
parms$theta_Si = parms$theta_Si * 10
parms$theta_Sn = parms$theta_Sn * 10
parms$theta_P = parms$theta_P * 1000
parms$probi = parms$probi * 10
parms$probn = parms$probi * 10
parms$f_Sn = parms$f_Sn * 10000
parms$f_Si = parms$f_Si * 10000
parms$f_In = parms$f_In * 1000
parms$f_Ii = parms$f_Ii * 1000000
parms$alpha = parms$alpha * 10000
parms$gamman = parms$gamman * 1e8
parms$gammai = parms$gammai * 1e4

parms$sigSi = parms$sigSi * 1e4
parms$sigIn = parms$sigIn * 1e5
parms$sigIi = parms$sigIi * 1e5
parms$sigF = parms$sigF * 10
parms$sigP = parms$sigP * 100
@

@@

$K_F$ & Alga carrying capacity&$10^2\cdot cells \cdot L ^{-1} $ & \Sexpr{mysignif(parms$K_F,2)}\\
$r_n$ & Native growth rate &$10^{-6} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$rn,2)}\\
$r_i$ & Invasive growth rate &$10^{-6} \cdot cells^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$ri,2)}\\
$\theta_{I_n}$ & Native infected host mortality rate&$ 10^{-1} \cdot day^{-1}$  & \Sexpr{mysignif(parms$theta_In,2)}\\
$\theta_{I_i}$ & Invasive infected host mortality rate&$10^{-1} \cdot day^{-1}$  & \Sexpr{mysignif(parms$theta_Ii,2)}\\
$\theta_{S_i}$ & Invasive susceptible host mortality rate&$10^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$theta_Si,2)}\\
$\theta_{S_n}$ & Native susceptible host mortality rate&$10^{-1} \cdot day^{-1}$ & \Sexpr{mysignif(parms$theta_Sn,2)}\\
$\theta_{P}$ & Spore degradation rate&$10^{-1} \cdot  day^{-3}$ & \Sexpr{mysignif(parms$theta_P,2)}\\
$p_n$ &Native probability of infection per spore &$10^{-4} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probn,2)}\\
$p_i$ & Invasive probability of infection per spore&$10^{-4} \cdot individuals \cdot spore^{-1} $ & \Sexpr{mysignif(parms$probi,2)}\\
$f_{S_n}$ & Native susceptible host filtering rate&$10^{-4} \cdot L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Sn,2)}\\
$f_{S_i}$ & Invasive susceptible host filtering rate&$10^{-4} \cdot L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_Si,2)} \\
$f_{I_n}$ & Native infected host filtering rate&$10^{-3} \cdot L \cdot day^{-1} \cdot individuals^{-1} $& \Sexpr{mysignif(parms$f_In,2)} \\
$f_{I_i}$ & Invasive infected host filtering rate&$10^{-6} \cdot L \cdot day^{-1} \cdot individuals^{-1} $ & \Sexpr{mysignif(parms$f_Ii,2)}\\

@@

$\alpha$ & Alga maximum exponential growth rate&$10^{-4} \cdot day ^{-1} $& \Sexpr{mysignif(parms$alpha,2)} \\
$\gamma_n$ & Native host consumption rate&$10^{-4} \cdot L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gamman,2)}\\
$\gamma_i$ & Invsaive host consumption rate&$10^{-4} \cdot L \cdot inidviduals^{-1} \cdot day ^{-1} $ & \Sexpr{mysignif(parms$gammai,2)}\\

@@

$\sigma_{S_i}$ & Standard deviation of brownian motion of susceptible invasive &$10^{-4} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigSi,2)}\\
$\sigma_{I_n}$ & Standard deviation of brownian motion of infected native&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIn,2)}\\
$\sigma_{I_i}$ & Standard deviation of brownian motion of infected invasive&$10^{-5} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigIi,2)}\\
$\sigma_{F}$ & Standard deviation of brownian motion of alga&$10^{-1} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigF,2)}\\
$\sigma_{P}$ &Standard deviation of brownian motion of parasite &$10^{-2} \cdot \sqrt{individual \cdot t^{-1}}$ & \Sexpr{mysignif(parms$sigP,2)}\\

@@

\begin{align}
&d S_n(t)=r_n S_n(t) dt- \theta_{S_n}S_n(t) dt -\delta S_n(t)dt + S_n(t) d\zeta_{S_n}, \\
&d S_i(t)=r_i S_i(t) dt- \theta_{S_i} S_i(t) dt -\delta S_i(t)dt + S_i(t) d\zeta_{S_i}, \\
&dF(t)=\alpha F(t)\left(1-\frac{F(t)}{K_F}\right)\cdot dt-\gamma_n F(t) S_n \cdot dt - \gamma_i F S_i \cdot dt -\delta F(t) \cdot dt+ \mu \cdot dt +F(t) d \zeta_F.\\
&\zeta_F \sim N \left( 0, \sigma_F^2dt\right) \\
&\zeta_{S_n} \sim N \left( 0, \sigma_{S_n}^2dt  \right) \\
&\zeta_{S_i} \sim N \left( 0, \sigma_{S_i}^2dt \right) 
\end{align}
